SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-07 docs/questioning.tex
add PL prefix
addressed PL comments

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it design-based}, {\it behavior-based} and {\it hybrid}.
The design-based category involves mechanisms that make applications
immune to different attack vectors by design~\cite{JL75,L81}. 
%% PL:
%% Explain what "by design" means.
The behavior-based category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
specific criteria~\cite{D76,A00}, that is, on specific behavioral
patterns. Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

Table~\ref{tab:avail} presents our findings regarding the availability
of the mechanism in terms of source code and corresponding
executables. We also examined the availability of the testbeds
mentioned in the paper. The check mark (\tick) indicates that at some
point within the publication, there is a homepage where the reader can
refer to, to download the corresponding software. {\bf AO} (Available
On-line) suggests that the software is available on-line but the
address was not mentioned in the paper, which probably indicates that
the authors made it available after the publication. The question mark
({\bf ?}), indicates that the a homepage for the software was included
in the publication but now it is not available.
%% PL: Do we need this paragraph here? Before we say that we tread
%% availability in the upcoming ``Analysis'' section.

@@

\subsection{Design-Based}
\label{sec:prot}

There exist three basic design-based approaches used to protect web
applications from executable source code injection attacks.

@@

As we will see in Subsection~\ref{sec:hybrid},
there are mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss} attacks.

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that utilize
this approach to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.
%% PL This paragraph repeats the material of the previous one, doesn't
%% it?

@@

This approach is used to detect attacks that target web users.
%% PL All of them more or less target web users, correct?
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
%% PL What kind of policies are these? Syntax-based, pattern matching?
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

@@

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (Locus {\sc x}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack~\cite{APKLM10} can be used to bypass
the above mechanisms. Such attacks share
many similarities with the {\it return-to-lib}
attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
When performing an attack like this the malicious user transfers the
execution to points that contain script code which in turn is used to
perform the attack.

@@

\subsection{Behavior-Based}

In this section we examine the two main behavior-based
approaches.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Locus {\sc z}).
However, this is a simplistic approach that may
lead to many false positives.
%% PL better to avoid characterisations and mention only the facts
%% (that it leads to many false positives, as shown by ...

It is possible to apply further checks when it is established that
tainted data have reached a sink (Locus {\sc z}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Locus {\sc z}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of specific features provided by {\sc
  php}.
%% PL Like what?
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).

A recent study showed that there are ways to circumvent the majority
of the above schemes~\cite{NBR14}. Furthermore, most of them are not
easy to deploy since the majority of input vectors, string operations
and output vectors of the application must be instrumented.

{\it Vogt et al.}~\cite{VFJKKV07} have developed a tainting scheme
that follows a different approach. Contrary to the aforementioned
schemes that operate on the server side, they track sensitive
information on the client side (Locus {\sc x}). This is a form of
positive taint tracking, where taint data is considered to be
legitimate. Their scheme detects JavaScript-driven {\sc xss} attacks
by ensuring that a script can send sensitive user data only to the
site from which it came from. This though, may lead to false alarms.
%% PL You mean a lot of false alarms?
{\it Stock et al.}~\cite{SLMS14} propose a scheme that also operates
in the browser (Locus {\sc x}). The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks. In such an attack, the malicious
payload is executed as a result of the modification of the {\sc dom}
environment so that the client-side code runs in an unanticipated
manner even if the {\sc http} response that triggers the injection
seems legitimate. The scheme is different from the previous one
because it marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript engine that
tracks the flow of attacker-controlled data. To detect potential
attacks the scheme utilizes {\sc html} and JavaScript parsers that can
identify the generation of code coming from tainted data.

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, consider an application that will send the
password for a forgetful user via email by executing the following
query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript attack
like the one mentioned in Subsubsection~\ref{sec:policy}.
%% PL Like the return-to-lib attack?

@@

{\sc xss-guard}~\cite{BV08} is a training scheme that also employs
parse-tree validation. During training, the scheme maps legitimate
scripts to {\sc http} responses. During production {\sc xss-guard}
retrieves for every script included in a response its parsed tree and
checks if it is one of those previously mapped to this response. Apart
from the comparison of the parsed trees, {\sc xss-guard} checks also
for an exact match of lexical entities. To achieve this, the scheme
utilizes that data structures of Firefox's JavaScript engine (Locus
{\sc x}). However, string literals are not compared literally and
therefore a false negative result could occur if an adversary performs
a mimicry attack~\cite{WS02}. For example, consider a banner rotator
that every time it runs, it creates a value that depends on the
current date and the length of the array that contains the references
of the various images to be displayed. Then, based on this value it
shows a specific image to a user. In a vulnerable web site that allows
users to post data and contains this banner rotator, a malicious user
could create and store a script that has the same code structure, with
the same JavaScript keywords like the above.
%% You mean like the rotator?
In this script the attacker could also include references to tiny
images hosted on a web server that is maintained by him in order to
retrieve the {\sc ip} addresses of the users that visit the vulnerable
site.

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking (see
Subsubsection~\ref{sec:taint}) together with parse-tree validation.
Diglossia was based on the theory of Ray and Ligatti~\cite{RL12b}
(also see Section~\ref{sec:attacks}) to detect {\sc dsl} code
injection attacks. In addition, it is actually the first, and so far
the only framework that detects No{\sc sql} injection attacks. When an
application computes an output string (query), Diglossia computes a
``shadow" of this string. Specifically, it maps all characters
introduced by the application to a shadow character set. This set does
not contain any characters coming from the tainted input. Then, the
scheme creates the tree of the query that is about to be executed and
compares it with the parsed tree of the ``shadow" (locus {\sc z}). If
the trees are not equal the application is probably under attack. Note
that Diglossia can be bypassed in the same way as other taint
tracking approaches~\cite{NBR14}.
%% PL which way is this exactly?

@@

% Comment 1: Regarding the frameworks that were
% not tested in terms of diagnostic performance
% at all, we see that they are mostly coming from
% the design-based category.

% Comment 2: Note though that implementations
% of design-based mechanisms can actually
% be bypassed.

@@

@@

Notes:
\begin{enumerate}
\item Note that most mechanisms that implement design-based
approaches have not been extensively tested in terms of
diagnostic performance.
\item We need to mention that there are mechanisms that
do not need to be established purely through testing
Since they provide systematic arguments as to why
the design is secure against attacks.
\item Policy enforcement is the
only approach that is
used to prevent {\sc csrf} attacks.
\item See j{\sc csrf} for {\sc soma}'s drawbacks.
\item Discuss about how tools can be easily modified
to detect other attacks ({\sc sql}rand).
\end{enumerate}

>>>>

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it etiological}, {\it symptomatic} and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
%% PL:
%% Explain what "by design" means.
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

@@

\subsection{Etiological}
\label{sec:prot}

There exist three basic etiological approaches used to protect web
applications from executable source code injection attacks.

@@

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.
%% PL This paragraph repeats the material of the previous one, doesn't
%% it?

@@

This approach is used to detect {\sc xss} and {\sc csrf} attacks.
%% PL All of them more or less target web users, correct?
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
Policies can be expressed through JavaScript extensions,
pattern matching or syntax specific settings.
%% PL What kind of policies are these? Syntax-based, pattern matching?
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

@@

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (Locus {\sc x}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack can be used to bypass the above mechanisms
as described extensively in reference~\cite{APKLM10}.
% Such attacks share
% many similarities with the {\it return-to-lib}
% attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
% When performing an attack like this the malicious user transfers the
% execution to points that contain script code which in turn is used to
% perform the attack.

@@

\subsection{Symptomatic}

In this section we examine the two main symptomatic
approaches.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Locus {\sc z}).
% However, this is a simplistic approach that may
% lead to many false positives.
%% PL better to avoid characterisations and mention only the facts
%% (that it leads to many false positives, as shown by ...

It is possible to apply further checks when it is established that
tainted data have reached a sink (Locus {\sc z}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Locus {\sc z}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
%% PL Like what?
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).

A recent study~\cite{NBR14} showed that there are ways to circumvent
the majority of the above schemes. Furthermore, most of them are not
easy to deploy since the majority of input vectors, string operations
and output vectors of the application must be instrumented.

{\it Vogt et al.}~\cite{VFJKKV07} have developed a tainting scheme
that follows a different approach. Contrary to the aforementioned
schemes that operate on the server side, they track sensitive
information on the client side (Locus {\sc x}). This is a form of
positive taint tracking, where taint data is considered to be
legitimate. Their scheme detects JavaScript-driven {\sc xss} attacks
by ensuring that a script can send sensitive user data only to the
site from which it came from.
% This though, may lead to false alarms.
%% PL You mean a lot of false alarms?
{\it Stock et al.}~\cite{SLMS14} propose a scheme that also operates
in the browser (Locus {\sc x}). The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks. In such an attack, the malicious
payload is executed as a result of the modification of the {\sc dom}
environment so that the client-side code runs in an unanticipated
manner even if the {\sc http} response that triggers the injection
seems legitimate. The scheme is different from the previous one
because it marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript engine that
tracks the flow of attacker-controlled data. To detect potential
attacks the scheme utilizes {\sc html} and JavaScript parsers that can
identify the generation of code coming from tainted data.

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} records all database
transactions without tagging each transaction with the
corresponding call-site. In this case a false negative result
is prominent. In particular, consider an application that will send the
password for a forgetful user via email by executing the following
query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\ref{sec:policy}.
%% PL Like the return-to-lib attack?

@@

{\sc xss-guard}~\cite{BV08} is a training scheme that also employs
parse-tree validation. During training, the scheme maps legitimate
scripts to {\sc http} responses. During production {\sc xss-guard}
retrieves for every script included in a response its parsed tree and
checks if it is one of those previously mapped to this response. Apart
from the comparison of the parsed trees, {\sc xss-guard} checks also
for an exact match of lexical entities. To achieve this, the scheme
utilizes that data structures of Firefox's JavaScript engine (Locus
{\sc x}). However, string literals are not compared literally and
therefore a false negative result could occur if an adversary performs
a mimicry attack~\cite{WS02}. For example, consider a banner rotator
that every time it runs, it creates a value that depends on the
current date and the length of the array that contains the references
of the various images to be displayed. Then, based on this value it
shows a specific image to a user. In a vulnerable web site that allows
users to post data and contains this banner rotator, a malicious user
could create and store a script that has the same code structure, with
the same JavaScript keywords contained in the rotator script.
%% PL You mean like the rotator?
In this script the attacker could also include references to tiny
images hosted on a web server that is maintained by him in order to
retrieve the {\sc ip} addresses of the users that visit the vulnerable
site.

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking (see
Subsubsection~\ref{sec:taint}) together with parse-tree validation.
Diglossia was based on the theory of Ray and Ligatti~\cite{RL12b}
(also see Section~\ref{sec:attacks}) to detect {\sc dsl} code
injection attacks. In addition, it is actually the first, and so far
the only framework that detects No{\sc sql} injection attacks. When an
application computes an output string (query), Diglossia computes a
``shadow" of this string. Specifically, it maps all characters
introduced by the application to a shadow character set. This set does
not contain any characters coming from the tainted input. Then, the
scheme creates the tree of the query that is about to be executed and
compares it with the parsed tree of the ``shadow" (locus {\sc z}). If
the trees are not equal the application is probably under attack. Note
that Diglossia can be bypassed in the same way as other taint
tracking approaches and is described in reference~\cite{NBR14}.
%% PL which way is this exactly?

@@

% Comment 1: Regarding the frameworks that were
% not tested in terms of diagnostic performance
% at all, we see that they are mostly coming from
% the etiological category.

% Comment 2: Note though that implementations
% of etiological mechanisms can actually
% be bypassed.

@@

Table~\ref{tab:avail} presents our findings regarding the availability
of the mechanism in terms of source code and corresponding
executables. We also examined the availability of the testbeds
mentioned in the paper. The check mark (\tick) indicates that at some
point within the publication, there is a homepage where the reader can
refer to, to download the corresponding software. {\bf AO} (Available
On-line) suggests that the software is available on-line but the
address was not mentioned in the paper, which probably indicates that
the authors made it available after the publication. The question mark
({\bf ?}), indicates that the a homepage for the software was included
in the publication but now it is not available.

@@

Notes:
\begin{enumerate}
\item Note that most mechanisms that implement etiological
approaches have not been extensively tested in terms of
diagnostic performance.
\item We need to mention that there are mechanisms that
do not need to be established purely through testing
Since they provide systematic arguments as to why
the design is secure against attacks.
\item Policy enforcement is the
only approach that is
used to prevent {\sc csrf} attacks.
\item See j{\sc csrf} for {\sc soma}'s drawbacks.
\item Discuss about how tools can be easily modified
to detect other attacks ({\sc sql}rand).
\end{enumerate}
